---
title: "Gun Control Chi-Square"
author: "Eleonore Durand"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This: I'm Eleonore and I'm updating this markdown file here.  

```{r}

library(tidyverse)
gc_df <- read_csv("gc_df.csv")


```


We are asking: Is there a significant difference in recommended gun control action for male and female respondents (should guns be easier, harder, or the same difficulty to get?)


```{r}
# Is it Thanksgiving yet? (Sam)

library(kableExtra)
```


Eleonore: data wrangling and counts 

```{r}


get_guns<- gc_df %>% 
  filter(gender != "NA", make_get_guns != "NA") %>% #filter out the NA values 
  count(gender, make_get_guns) %>% #for each level of the gender column, counts responses
  spread(make_get_guns, n) %>%  # by 'n'
  select(-gender) #using '-' tells it to remove the column

rownames(get_guns)<- c ("Female", "Male")


```

My name is Sam the collaborator and I'm going to be doing chi-square test for independence!

We ask: "Is there a significant association between sex and opinion on whether guns should be more difficult, easier, or the same difficulty to get?"

```{r}

#Chi-sqaure compares proportions of counts. What are the actual proportions? How do each differ?

#First, look at the actual proportions.

guns_prop <- prop.table(as.matrix(get_guns), 1) #updating class of make_guns so it works with prop.table function. # ,1) tells it to use the entire row (i.e. female) to find proportion in row?

guns_prop

#Proportions are something you would want to report!

# Runn chi-square test for indepence. See above test question outside of the code. Make sure you made a table that is in the correct form before--then the test is simple!

guns_x2 <- chisq.test(get_guns)
guns_x2

# H0: there is no difference in proportion of M:F respondents in what they think about gun access
# HA: there is a sign diff in proportion of M:F respondents in what they think about gun access

"Recommended government action on gun access differed significantly between male (n= ) and female (n= ) respondents ($\chi^2$(`r guns_x2$parameter`) = ...., p < 0.001. Notable, a much higher proportion of female respondents (57.4%)  recommend making it more difficult to get guns, compared to 40.4% for male respondents.

Could follow up with pairwise.

# guns_x2$stdres counts are different enough to derive sig result
```

